---
title: "Preliminary EDA"
author: "Tarun Sudhams"
date: "11/17/2018"
output: html_document
---

Let's quickly explore the dimensionality and the behaviors of different categorical features present in the dataset. But before that we will try to import the required libraries for our study.

```{r}
library(tidyverse)
library(grid)
library(gridExtra)
```

Assigning train and test data to variables.

```{r}
train = read_csv('train.csv')
test = read_csv('test.csv')
```

A quick 
```{r}
head(train)
```

```{r}
train$Target[train$Target == 1] = 'Extreme Poverty'
train$Target[train$Target == 2] = 'Moderate Poverty'
train$Target[train$Target == 3] = 'Vulnerable Household'
train$Target[train$Target == 4] = 'Non-Vulnerable Household'

train %>%
ggplot(aes(Target))+
geom_bar(color = 'black', fill = 'tomato')+
xlab("Target Classes")+
ylab("Target Classes Count")

```

``` {r}
train %>%
ggplot(aes(x = as.factor(v18q)))+
geom_bar(colour = "black", fill = "blue")+
facet_wrap(~Target, scales = "free", ncol = 3)+ 
theme(axis.text.x = element_text(hjust = 1, size = 11))+
theme(strip.text = element_text(size = 10, face = "bold"))+
labs(x = "Tablet Owner", y = "Number of Households")+
scale_x_discrete(labels = c("No", "Yes"))

```
Majority of the households do not own a tablet but many do. And the interesting thing is that people in extreme poverty households own a tablet! Lets look at how they fare when compared together.
``` {r}
train %>%
ggplot(aes(as.factor(v18q)))+
geom_bar(aes(fill = as.factor(Target)), position = "dodge", color = "grey")+
labs(x = "Tablet Owner", y = "Number of Households")+
guides(fill = guide_legend('Household Type'))+
theme(axis.text.x = element_text(hjust = 1, size = 10))+
scale_x_discrete(labels = c("No", "Yes"))
```
Now lets look at the number of tablets owned by households which do own a tablet

``` {r}
train %>%
filter(!is.na(v18q1)) %>%
ggplot(aes(as.factor(v18q1)))+
geom_bar(color = "blue")

```
So the maximum number of tablets owned is 1 but households do own 6 tablets

``` {r}

train %>%
filter(!is.na(v18q1)) %>%
ggplot(aes(x = as.factor(v18q1)))+
geom_bar(colour = "black", fill = "green")+
facet_wrap(~Target, scales = "free", ncol = 3)+ 
theme(axis.text.x = element_text(hjust = 1, size = 10))+
theme(strip.text = element_text(size = 9, face = "bold"))+
labs(x = "Number of tablets owned", y = "Number of Households")

```


```{r}
train %>%
filter(!is.na(v18q1)) %>%
ggplot(aes(as.factor(v18q1)))+
geom_bar(aes(fill = as.factor(Target)), position = "dodge", color = "black")+
labs(x = "Owns a tablet or not?", y = "Count")+
guides(fill = guide_legend('Household Type'))+
theme(axis.text.x = element_text(hjust = 1, size = 9))

```


## Male Distribution
```{r}
train %>%
ggplot(aes(as.factor(r4h1)))+
geom_bar(color = "grey", fill = "red", alpha = 0.5)+
labs(x = "Males younger than 12", "Number of Households")
```

```{r}
train %>%
ggplot(aes(as.factor(r4h2)))+
geom_bar(color = "grey", fill = "black", alpha = 0.6)+
labs(x = "Males older than 12", "Number of Households")
```

```{r}
train %>%
ggplot(aes(x = as.factor(Target)))+
geom_bar(colour = "black", fill = "red", alpha = 0.6)+
facet_wrap(~as.factor(r4h1), scales = "free", ncol = 3)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))+
theme(strip.text = element_text(size = 9, face = "bold"))+
labs(title = "Association between Household Type and number of males younger than 12", x = "Household Type", y = "Number of Households")
```
We see that the number of males younger than 12 are high in extreme poverty and non-vulnerable households.

```{r}
train %>%
ggplot(aes(x = as.factor(Target)))+
geom_bar(colour = "grey19", fill = "maroon", alpha = 0.6)+
facet_wrap(~as.factor(r4h2), scales = "free", ncol = 3)+ 
theme(axis.text.x = element_text(angle = 15, hjust = 1, size = 7))+
theme(strip.text = element_text(size = 9, face = "bold"))+
labs(title = "Association between Household Type and number of males older than 12 ", x = "Household Type", y = "Number of Households")
```
The number of data points concerning non-vulnerable households is very large. We can remove some datapoints and analyze them
to understand the relationship of low income households and compare them

```{r}
train %>%
filter(Target != "Non-Vulnerable Household") %>%
ggplot(aes(as.factor(r4h2)))+
geom_bar(aes(fill = as.factor(Target)), position = "dodge", color = "grey", alpha = 0.5)+
labs(x = "Number of Males older than 12", y = "Number of Households")+
guides(fill = guide_legend('Household Type'))+
theme(axis.text.x = element_text(hjust = 1, size = 9))
```
This reveals an interesting property about the extreme poverty household. There is a high population of males younger than 12 in extreme poverty households 
while the concentration of males older than 12 is lesser.

```{r}
train %>%
ggplot(aes(as.factor(r4h3)))+
geom_bar(color = "black", fill = "gold", alpha = 0.6)+
labs(x = "Total Number of Males in Household", "Number of Households")
```
```{r}
train %>%
ggplot(aes(x = as.factor(r4h3)))+
geom_bar(colour = "black", fill = "red", alpha = 0.5)+
facet_wrap(~as.factor(Target), scales = "free", ncol = 3)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))+
theme(strip.text = element_text(face = "bold",size = 10 ))+
labs(title = "Association between Household Type and total number of males", x = "Type of Household", y = "Number of Households")

```
Female Distribution 

```{r}
train %>%
ggplot(aes(as.factor(r4m1)))+
geom_bar(color = "black", fill = "red", alpha = 0.5)+
labs(x = "Females younger than 12", "Number of Households")
```

```{r}
train %>%
ggplot(aes(as.factor(r4m2)))+
geom_bar(color = "black", fill = "blue", alpha = 0.5)+
labs(x = "Females older than 12", "Number of Households")
```
```{r}
train %>%
ggplot(aes(x = as.factor(Target)))+
geom_bar(colour = "black", fill = "red", alpha = 0.6)+
facet_wrap(~as.factor(r4m1), scales = "free", ncol = 3)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))+
theme(strip.text = element_text(size = 9, face = "bold"))+
labs(title = "Association between Household Type and number of females younger than 12", x = "Household Type", y = "Number of Households")
```
```{r}
train %>%
ggplot(aes(x = as.factor(Target)))+
geom_bar(colour = "grey19", fill = "maroon", alpha = 0.6)+
facet_wrap(~as.factor(r4m2), scales = "free", ncol = 3)+ 
theme(axis.text.x = element_text(angle = 15, hjust = 1, size = 7))+
theme(strip.text = element_text(size = 9, face = "bold"))+
labs(title = "Association between Household Type and number of females older than 12", x = "Household Type", y = "Number of Households")
```
```{r}
train %>%
ggplot()+
geom_bar(aes(x = as.factor(r4m1)), colour = "black", fill = "red", alpha = 0.5)+
geom_bar(aes(x = as.factor(r4m2)), colour = "black", fill = "blue", alpha = 0.25)+
facet_wrap(~as.factor(Target), scales = "free", ncol = 3)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))+
theme(strip.text = element_text(size = 9, face = "bold"))+
labs(title = "Association between Household Type and number of females of different ages", x = "Type of Household", y = "Number of Households")
```
There is not a huge population of females younger than 12 in majority of the households. However, the number of females older
than 12 are 1 in a majority of the households with the number going upto 5 in households with extreme poverty

```{r}
train %>%
filter(Target != "Non-Vulnerable Household") %>%
ggplot(aes(as.factor(r4m2)))+
geom_bar(aes(fill = as.factor(Target)), position = "dodge", color = "grey", alpha = 0.5)+
labs(x = "Females older than 12", y = "Number of Households")+
guides(fill = guide_legend('Household Type'))+
theme(axis.text.x = element_text(hjust = 1, size = 9))
```
```{r}
train %>%
filter(Target != "Non-Vulnerable Household") %>%
ggplot(aes(as.factor(r4m1)))+
geom_bar(aes(fill = as.factor(Target)), position = "dodge", color = "black", alpha = 0.6)+
labs(x = "Females younger than 12", y = "Number of Households")+
guides(fill = guide_legend('Household Type'))+
theme(axis.text.x = element_text(hjust = 1, size = 9))
```
Females younger than 12 seem to be present majorly in moderate poverty households

```{r}
train %>%
ggplot(aes(as.factor(r4m3)))+
geom_bar(color = "grey", fill = "red", alpha = 0.6)+
labs(x = "Total Number of Females in the Household", "Number of Households")

```

```{r}
train %>%
ggplot(aes(x = as.factor(r4m3)))+
geom_bar(colour = "black", fill = "blue", alpha = 0.5)+
facet_wrap(~as.factor(Target), scales = "free", ncol = 3)+ 
theme(axis.text.x = element_text(angle = 45, size = 8, hjust = 1))+
theme(strip.text = element_text(size = 9, face = "bold"))+
labs(title = "Association between Type of Household and total number of females ", x = "Household Type", y = "Number of Households")
```
8. Person Distribution
```{r}
train %>%
ggplot(aes(as.factor(r4t1)))+
geom_bar(color = "grey", fill = "red", alpha = 0.6)+
labs(x = "Persons younger than 12", "Number of Households")
```
```{r}
train %>%
ggplot(aes(as.factor(r4t2)))+
geom_bar(color = "grey", fill = "blue", alpha = 0.5)+
labs(x = "Persons older than 12", "Number of Households")

```
```{r}
train %>%
ggplot()+
geom_bar(aes(as.factor(r4t1)), colour = "black", fill = "green", alpha = 0.5)+
facet_wrap(~as.factor(Target), scales = "free", ncol = 3)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))+
theme(strip.text = element_text(size = 9, face = "bold"))+
labs(title = "Association between Type of Households and number of persons younger than 12 ", x = "Household Type", y = "Number of Households")
```
Moderate poverty and extreme poverty households have a huge number of persons younger than 12 whereas vulnerable and non-vulnerable households have a lesser number
of persons age less than 12

```{r}
train %>%
ggplot()+
geom_bar(aes(as.factor(r4t2)), colour = "black", fill = "green", alpha = 0.5)+
facet_wrap(~as.factor(Target), scales = "free", ncol = 3)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))+
theme(strip.text = element_text(size = 9, face = "bold"))+
labs(title = "Association between Type of Households and number of persons older than 12 ", x = "Household Type", y = "Number of Households")
```
There is a reversal of results here with the number of persons older than 12 being higher in non-vulnerable and vulnerable households.

```{r}
train %>%
mutate(
    mean = r4t3/rooms
) %>%
ggplot(aes(x = mean))+
geom_histogram(colour = "black", fill = "blue", bins = 30, alpha = 0.4)+ 
facet_wrap(~Target, scales = "free", ncol = 3)+ 
theme(axis.text.x = element_text(hjust = 1, size = 12))+
theme(strip.text = element_text(face = "bold", size = 10))+
labs(title =" Mean Number of Persons vs Rooms", x = "Mean number of persons", y = "Number of Households", size = 12)
```
The number of persons in a room goes up as the households get poorer.

9. Number of Years of Schooling
```{r}
train %>%
filter(Target == "Extreme Poverty") %>%
ggplot(aes(x = as.factor(Target), stat = 'bin', fill = as.factor(escolari)))+
geom_bar(position = 'fill', color = 'black')+
coord_polar("y")+
labs(x = "", y = "")+
guides(fill = guide_legend("Years of schooling"))
```
```{r}
train %>%
filter(Target == "Moderate Poverty") %>%
ggplot(aes(x = as.factor(Target), stat = 'bin', fill = as.factor(escolari)))+
geom_bar(position = 'fill', color = 'black')+
coord_polar("y")+
labs(x = "", y = "")+
guides(fill = guide_legend("Years of schooling"))
```
```{r}
train %>%
filter(Target == "Vulnerable Household") %>%
ggplot(aes(x = as.factor(Target), stat = 'bin', fill = as.factor(escolari)))+
geom_bar(position = 'fill', color = 'black')+
coord_polar("y")+
labs(x = "", y = "")+
guides(fill = guide_legend("Years of schooling"))
```
```{r}
train %>%
filter(Target == "Non-Vulnerable Household") %>%
ggplot(aes(x = as.factor(Target), stat = 'bin', fill = as.factor(escolari)))+
geom_bar(position = 'fill', color = 'black')+
coord_polar("y")+
labs(x = "", y = "")+
guides(fill = guide_legend("Years of schooling"))
```

Majority of persons in extreme poverty households have 0 years of households
